Clustering Constraints on the Relative Sizes of Central and Satellite Galaxies

### 2017-11-26 paper/galsize_paper.tex
Minor changes to Section 8
Minor edits to captions

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=9cm]{FIGS/sdss_size_vs_mstar_scatter.pdf}
\includegraphics[width=\textwidth]{FIGS/sdss_small_large_sample_definitions.pdf}
\caption{
{\bf Definition of ``small" and ``large" galaxies.} For a volume-limited SDSS galaxy sample defined by $M_{\ast}>10^{10.25}M_{\odot},$ the {\em top panel} visually demonstrates how we classify galaxies into ``small" and ``large" subsamples. As described in detail in \S\ref{subsec:sizedef}, we compute the median value $\median{\rhalf}{\mstar}$ at the stellar mass of each galaxy in the sample, and divide the sample into two based on this $\mstar-$dependent cut. The {\em bottom left panel} compares the stellar mass functions $\phi(\mstar)$ of the two samples, confirming that our method for separating ``small" from ``large" galaxies yields subsamples with statistically indistinguishable $\phi(\mstar)$. The {\em bottom right panel} shows histograms of $\rhalf$ for the two subsamples, which partially overlap due to the stellar mass range spanned by the sample.
}
\label{fig:sizedefinition}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=11cm]{FIGS/rvir_only_wp_large_small_absolute.pdf}
\caption{
{\bf $\rhalf-$dependence of galaxy clustering.}
Red and blue points with error bars show our SDSS measurements of the clustering of small and large galaxies, respectively. For each volume-limited sample of $\mstar-$complete galaxies, the small and large subsamples have identical stellar mass functions, as shown in Figure \ref{fig:sizedefinition}. Small galaxies cluster much more strongly relative to large galaxies of the same stellar mass. Solid curves show the clustering predictions of the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}. The $\rvir-$based model inherits the shortcoming of ordinary abundance matching shown in Figure \ref{fig:baseline_sham_clustering}, although the model faithfully captures the {\em relative} clustering of small vs. large galaxies, as shown in Figure \ref{fig:clustering_ratio_upshot}.
}
\label{fig:rvir_only_clustering_absolute}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=11cm]{FIGS/rvir_only_wp_ratios.pdf}
\caption{
{\bf $\rhalf-$dependence of galaxy clustering.}
Closely related to Figure \ref{fig:rvir_only_clustering_absolute}, the y-axes show {\em clustering strength ratios,} defined as $\delta\wproj/\wproj\equiv(w_{\rm p}^{\rm large} - w_{\rm p}^{\rm small})/w_{\rm p}^{\rm all}.$ Thus a y-axis value of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of the same stellar mass. Solid curves show the clustering ratio predictions of the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}. Normalizing the measurements and predictions by $w_{\rm p}^{\rm all}$ scales away the shortcoming of ordinary abundance matching (see Figure \ref{fig:baseline_sham_clustering}), highlighting the successful prediction of the $\rvir-$based model for the $\rhalf-$dependence of galaxy clustering.
}
\label{fig:clustering_ratio_upshot}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm9p75_sm10p25.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction2.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p25_sm10p75.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction3.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p75_sm11p25.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, in the {\em left panels} we show predictions for future measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the observed trends are reversed and weaker than those predicted here: for both blue and red samples, the observed $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show that $\Delta\Sigma$ of small galaxies is significantly stronger relative to large galaxies of the same stellar mass $\mstar\lesssim10^{11}\msun.$ We additionally predict that at sufficiently large mass $\mstar\gtrsim10^{11}\msun,$ the trend of small-scale $\Delta\Sigma$ with $\rhalf$ reverses sign. The {\em right panels} display the physical explanation for this prediction. For $\mstar\lesssim10^{11}\msun,$ satellites preferentially populating the small sample heavily weight high-mass halos, boosting the lensing signal on all scales. For $\mstar\gtrsim10^{11}\msun,$ when $F_{\rm sat}\lesssim10\%,$ central galaxies dominate the signal, and the $\rvir-$based model predicts that $\rhalf$ increases with $\mhalo$ for centrals. See text for further details. 
}
\label{fig:lensingprediction}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

Our results provide an integral constraint for more complex and fine-grained models of galaxy size, such as semi-analytic models and hydrodynamical simulations. For convenience, Figure \ref{fig:censatsizeratios} provides a simple summary statistic that can be used as a diagnostic for calibrating alternative modeling efforts. We view the present work as a pilot study that motivates a Bayesian inference program to tightly constrain the galaxy size-halo connection with forward modeling techniques, in direct analogy to the literature on the stellar-to-halo-mass relation.

>>>>

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=9cm]{FIGS/sdss_size_vs_mstar_scatter.pdf}
\includegraphics[width=\textwidth]{FIGS/sdss_small_large_sample_definitions.pdf}
\caption{
{\bf Definition of ``small" and ``large" galaxies.} For a volume-limited SDSS galaxy sample defined by $M_{\ast}>10^{10.25}M_{\odot},$ the {\em top panel} visually demonstrates how we classify galaxies into ``small" and ``large" subsamples in an $\mstar$-dependent manner. As described in detail in \S\ref{subsec:sizedef}, we compute the median value $\median{\rhalf}{\mstar}$ at the stellar mass of each galaxy in the sample, and divide the sample into two based on this $\mstar-$dependent cut. The {\em bottom left panel} compares the stellar mass functions $\phi(\mstar)$ of the two samples, confirming that our method for separating ``small" from ``large" galaxies yields subsamples with statistically indistinguishable $\phi(\mstar)$. The {\em bottom right panel} shows histograms of $\rhalf$ for the two subsamples, which partially overlap due to the stellar mass range spanned by the sample.
}
\label{fig:sizedefinition}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=11cm]{FIGS/rvir_only_wp_large_small_absolute.pdf}
\caption{
{\bf $\rhalf-$dependence of galaxy clustering: comparison to $\rvir$-based model.}
Red and blue points with error bars show our SDSS measurements of the clustering of small and large galaxies, respectively. For each volume-limited sample of $\mstar-$complete galaxies, the small and large subsamples have identical stellar mass functions, as shown in Figure \ref{fig:sizedefinition}. Small galaxies cluster much more strongly relative to large galaxies of the same stellar mass. Solid curves show the clustering predictions of the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}. The $\rvir-$based model inherits the shortcoming of ordinary abundance matching shown in Figure \ref{fig:baseline_sham_clustering}, although the model faithfully captures the {\em relative} clustering of small vs. large galaxies, as shown in Figure \ref{fig:clustering_ratio_upshot}.
}
\label{fig:rvir_only_clustering_absolute}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=11cm]{FIGS/rvir_only_wp_ratios.pdf}
\caption{
{\bf $\rhalf-$dependence of galaxy clustering: comparison to $\rvir$-based model.}
Closely related to Figure \ref{fig:rvir_only_clustering_absolute}, y-axes show {\em clustering strength ratios,} defined as $\delta\wproj/\wproj\equiv(w_{\rm p}^{\rm large} - w_{\rm p}^{\rm small})/w_{\rm p}^{\rm all}.$ Thus a y-axis value of $-0.5$ corresponds to small galaxies being $50\%$ more strongly clustered than large galaxies of the same stellar mass. Solid curves show predictions of the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}. Normalizing the measurements and predictions by $w_{\rm p}^{\rm all}$ scales away the shortcoming of ordinary abundance matching (see Figure \ref{fig:baseline_sham_clustering}), highlighting the successful prediction of the $\rvir-$based model for the $\rhalf-$dependence of galaxy clustering.
}
\label{fig:clustering_ratio_upshot}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

%---------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm9p75_sm10p25.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction2.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p25_sm10p75.pdf}
\includegraphics[width=8cm]{FIGS/rvir_only_lensing_prediction3.pdf}
\includegraphics[width=8cm]{FIGS/host_mass_distributions_sm10p75_sm11p25.pdf}
\caption{
{\bf Prediction for $\rhalf-$dependence of galaxy lensing.}
Using the $\rvir-$based model described in \S\ref{subsubsec:rvirmodel}, in the {\em left panels} we show predictions for future measurements of the $\rhalf-$dependence of galaxy lensing of $\mstar-$complete samples, where ``small" and ``large" subsamples are defined as in \S\ref{subsec:sizedef}. To date, the $\rhalf-$dependence of $\Delta\Sigma$ has only been measured for color-selected samples \citep{charlton_etal17}, in which the observed trends are reversed and weaker than those predicted here: for both blue and red samples, the observed $\Delta\Sigma$ of large galaxies is (slightly) stronger relative to small galaxies. As a generic consequence of satellites being smaller than centrals of the same mass, we predict that future lensing measurements of $\mstar-$complete samples will show that $\Delta\Sigma$ of small galaxies is significantly stronger relative to large galaxies of the same stellar mass $\mstar\lesssim10^{11}\msun.$ We additionally predict that at sufficiently large mass $\mstar\gtrsim10^{11}\msun,$ the trend of small-scale $\Delta\Sigma$ with $\rhalf$ reverses sign. The {\em right panels} display the physical explanation for this prediction. For $\mstar\lesssim10^{11}\msun,$ satellites preferentially populating the small sample heavily weight high-mass halos, boosting the lensing signal on all scales. For $\mstar\gtrsim10^{11}\msun,$ when $F_{\rm sat}\lesssim10\%,$ central galaxies dominate the signal, and the $\rvir-$based model predicts that $\rhalf$ increases with $\mhalo$ for centrals. See \S\ref{sec:previous_work} for further details. 
}
\label{fig:lensingprediction}
\end{figure*}
%-----------------------------------------------------------------------------------------------------

@@

We view the present work as a pilot study that motivates a Bayesian inference program to tightly constrain the galaxy size-halo connection with forward modeling techniques, in direct analogy to the literature on the stellar-to-halo-mass relation.
Our results provide an integral constraint for more complex and fine-grained models of galaxy size, such as semi-analytic models and hydrodynamical simulations. For convenience, Figure \ref{fig:censatsizeratios} provides a simple summary statistic that can be used as a diagnostic for calibrating alternative modeling efforts.
